Analytical choices for analyzing multidimensional behavior - Many analysts test hypotheses about human speech.

### 2021-03-10 RR_manuscript/Draft_RR.Rmd
:memo: - add comments

Data analysis comes with many decisions, like how to measure a given phenomenon or behavior, which data to submit to statistical modeling and which to exclude in the final analysis, or what inferential decision making procedure to apply.
This can be problematic because humans show cognitive biases that can lead to erroneous inferences.
Humans filter the world in irrational ways [e.g., @tversky1974judgment], seeing coherent patterns in randomness [@brugger2001], convincing themselves of the validity of prior expectations ["I knew it", @nickerson1998confirmation], and perceiving events as being plausible in hindsight ["I knew it all along", @fischhoff1975hindsight].
In conjunction with an academic incentive system that rewards certain discovery processes more than others [@sterling1959publication; @koole2012rewarding], we often find ourselves exploring many possible analytical pipelines, but only reporting a select few. <!-- in connection with = referring to; in conjunction with = together, coupled with. I think we want the later, but change if I'm mistaken (JVC) --> 
This issue is particularly amplified in fields in which the raw data lend themselves to many possible ways of being measured [@roettger2019researcher].
Combined with a wide variety of methodological and theoretical traditions as well as varying levels of statistical training across subfields, the inherent flexibility of data analysis might lead to a vast plurality of analytic approaches that can lead to different scientific conclusions.
Consequently, there might be many published papers that present overconfident interpretations of their data based on idiosyncratic analytic strategies [e.g., @simmons2011false; @gelman2014statistical].
These interpretations are either associated with an unknown amount of uncertainty or lend themselves to alternative interpretation if analyzed differently.
However, instead of being critically evaluated, scientific results often remain unchallenged in the publication record.
Despite recent efforts to improve transparency and reproducibility [e.g. @miguel2014promoting; @klein2018practical] and the advent of freely available and accessible infrastructures, such as those provided by the Open Science Framework (osf.io), critical re-analyses of published analytic strategies are still uncommon because data sharing remains rare [@wicherts2006poor].

Analytical flexibility has been widely discussed from a conceptual point of view [@simmons2011false; @wagenmakers2012agenda; @nosek2014method] and in regard to its application in individual scientific fields [e.g. Wichert et al. 2015, Charles et al. 2019, @roettger2019researcher], notwithstanding, there are still many unknowns regarding the extent of analytical plurality in practice.
<!-- TODO: add refs for wichert et al. 2015, Charles et al. 2019... are we certain this isn't @wicherts2016? (JVC) -->
Recent collaborative attempts have started to shed light on how different analysts tackle the same data set and have revealed a large amount of variability. <!-- I think this should be part of the previous paragraph. -->

@@

Speech categories are inherently multidimensional and dynamic: they consist of clusters of parameters that are modulated over time.
The acoustic parameters of one category are usually asynchronous, i.e. they appear at different time points in the unfolding signal, and overlap with parameters of other categories [e.g. @jongman2000acoustic; @lisker1986voicing; @summerfield1984; @winter2014spoken].
<!-- TODO: add summerfield ref-->
A classical example is the distinction between voiced and voiceless stops in English (i.e. /b/ and /p/ in *bear* vs *pear*).
This lexical contrast is manifested by many acoustic features which can differ depending on several factors, such as position of the consonant in the word and surrounding sounds [@lisker1977rapid]. <!-- why not "phonological contrast"? --> 
Furthermore, correlates of the contrast can even be found away from the consonant, in temporally distant speech units.
For example, the initial /l/ of the English words *led* and *let* is affected by the voicing of the final consonant (/t, d/) [@hawkins2004influence].

@@

When looking at phrase-level temporal domains, the number of possible analytical pipelines increases exponentially. <!-- this is hyperbole, but seems (possibly) more appropriate than "explodes" -->
This plurality of analytical paths is illustrated in Figure \@ref(fig:forkingPaths).
<!-- When comparing two utterances such as (2a) and (2b), there are many things to consider.  (I think we can do without this sentence.) -->
Even if we have decided to compare fundamental frequency (f0) of only the word *another* across the two utterances, there are still many choices to be made, all of which need to be justified.
For example, we could measure f0 at specific points in time like the onset of the window, the offset, or the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarize f0 across the entire window and extract the mean, median or standard deviation of f0, all of which have been used to analyze speech data [see @gordon2017acoustic].
And the garden of forking paths does not stop here.
<!-- TODO: add examples of pitch tracking options -->
<!-- NOTE: (TR) is the figure better now? -->
Figure \@ref(fig:forkingPaths), illustrates a specific option to automatically calculate f0.
Knowing that the signals and the measures thereof can be noisy, another option would be to smooth them using one algorithm and degree over others, or remove values that substantially deviate from surrounding values or expectations, either manually or automatically, and so on. <!-- Im having a bit of trouble with this sentence. I tried to fix it and now I am realizing I don't quite follow (JVC) -->

@@

Even looking at morphosyntactically identical expressions, speakers can modulate utterances via suprasegmental acoustic properties like temporal and spectral modifications of the segments involved [e.g., @ladd2008intonational].
Most prominently, languages use intonation to signal discourse relationships between referents (among other functions).
Intonation marks discourse-relevant referents for being new or given information to guide the listeners’ interpretation of incoming messages. 
Beyond structuring information relative to the discourse, a few studies suggested that speakers might use intonation to signal atypical lexical combinations [e.g. dimitrova2008prosodic; dimitrova2009did]. 
Referential expressions such as *blue banana* were produced with greater prosodic prominence than more typical referent such as *yellow banana*.
If true, these findings are in line with the idea of resource-efficient, rational language users who modulate their speech in order to facilitate listeners' comprehension process.
However, the original conclusions of above studies were based on small (n=10), unjustified sample sizes. 
Moreover, the data is not available, and the statistical analysis used anti-conservative co-variance structures, leaving reason to doubt the original conclusions. <!-- "lending doubt on" sounded a bit weird to me (JVC) -->

@@

<!-- TODO: add materials link and mailing lists -->
The initiating authors (SC, JC, TR) created an online landing page providing a general description of the project [https://many-speech-analyses.github.io](https://many-speech-analyses.github.io) and a short pre-recorded slide show that summarizes the study and research question (LINK). <!-- same as above, or OSF -->
The project will be advertised via social media, using mailing lists for linguistic and psychological societies, and via word of mouth.
The target population comprises active speech science researchers with a graduate/doctoral degree (or currently studying for a graduate/doctoral degree) in relevant disciplines.
Researchers can choose to work independently or in a small team.
For the sake of simplicity, we will refer to single researcher or small teams as "analysis teams".

@@

<!-- I understand "posterior probability" to refer to a point estimate, i.e., the posterior probability that B > 0 = some value between 0 and 1. I may be mistaken, but I think you are referring to the posterior distribution in the above paragraph -->

@@

<!-- Note: I keep changing meta-analytical to meta-analytic (the -cal sounds weird to me). My unscientific google searching shows that people use both, but meta-analytic is much more common (nearly 4 million hits vs. 500k) -->

@@

### Analytical and demographic predictors affecting effect sizes {#ana-factors}
<!-- TR: I think we should use the term predictors. Factors is really ambiguous -->
As a second step, we will investigate (a) the extent to which the individual standardized effect sizes are affected by a series of predictors related to analytical and demographic factors (see below); and (b) the extent to which deviations from the meta-analytic point estimate by the individual standardized effect sizes relate to those same predictors.
We will fit two Bayesian regression models, one for (a) and one for (b).
<!-- TODO: check if stand. effect sizes and deviations are equivalent. -->
with each team's group-level coefficient $\alpha_{\text{t}[i]}$ as the outcome variable, and the analytical and demographic factors described below as predictors.
The coefficient $\alpha_{\text{t}[i]}$ can be interpreted as the deviation score of the team's $\eta_i$ from the meta-analytic point estimate, allowing us to determine potential effects of analytical and demographic factors on how the teams deviate from the meta-analytic mean.
Since each $\alpha_{\text{t}[i]}$ comes with uncertainty, quantified by its standard deviation $\sigma_{\alpha_{\text{t}[i]}}$, we will fit a measurement-error model, as we did in the meta-analytic model above.

>>>>

Data analysis comes with many decisions, like how to measure a given phenomenon or behavior, which data to submit to statistical modeling and which to exclude in the final analysis, or what inferential decision making procedure to apply.
This can be problematic because humans show cognitive biases that can lead to erroneous inferences.
Humans filter the world in irrational ways [e.g., @tversky1974judgment], seeing coherent patterns in randomness [@brugger2001], convincing themselves of the validity of prior expectations ["I knew it", @nickerson1998confirmation], and perceiving events as being plausible in hindsight ["I knew it all along", @fischhoff1975hindsight].
In conjunction with an academic incentive system that rewards certain discovery processes more than others [@sterling1959publication; @koole2012rewarding], we often find ourselves exploring many possible analytical pipelines, but only reporting a select few.
<!-- NOTE: (JVC) in connection with = referring to; in conjunction with = together, coupled with. I think we want the later, but change if I'm mistaken -->
This issue is particularly amplified in fields in which the raw data lend themselves to many possible ways of being measured [@roettger2019researcher].
Combined with a wide variety of methodological and theoretical traditions as well as varying levels of statistical training across subfields, the inherent flexibility of data analysis might lead to a vast plurality of analytic approaches that can lead to different scientific conclusions.
Consequently, there might be many published papers that present overconfident interpretations of their data based on idiosyncratic analytic strategies [e.g., @simmons2011false; @gelman2014statistical].
These interpretations are either associated with an unknown amount of uncertainty or lend themselves to alternative interpretation if analyzed differently.
However, instead of being critically evaluated, scientific results often remain unchallenged in the publication record.
Despite recent efforts to improve transparency and reproducibility [e.g. @miguel2014promoting; @klein2018practical] and the advent of freely available and accessible infrastructures, such as those provided by the Open Science Framework (osf.io), critical re-analyses of published analytic strategies are still uncommon because data sharing remains rare [@wicherts2006poor].

Analytical flexibility has been widely discussed from a conceptual point of view [@simmons2011false; @wagenmakers2012agenda; @nosek2014method] and in regard to its application in individual scientific fields [e.g. Wichert et al. 2015, Charles et al. 2019, @roettger2019researcher], notwithstanding, there are still many unknowns regarding the extent of analytical plurality in practice.
<!-- TODO: add refs for wichert et al. 2015, Charles et al. 2019... are we certain this isn't @wicherts2016? (JVC) -->
Recent collaborative attempts have started to shed light on how different analysts tackle the same data set and have revealed a large amount of variability.
<!-- NOTE: (?) I think this should be part of the previous paragraph. -->

@@

Speech categories are inherently multidimensional and dynamic: they consist of clusters of parameters that are modulated over time.
The acoustic parameters of one category are usually asynchronous, i.e. they appear at different time points in the unfolding signal, and overlap with parameters of other categories [e.g. @jongman2000acoustic; @lisker1986voicing; @summerfield1984; @winter2014spoken].
<!-- TODO: add summerfield ref-->
A classical example is the distinction between voiced and voiceless stops in English (i.e. /b/ and /p/ in *bear* vs *pear*).
This lexical contrast is manifested by many acoustic features which can differ depending on several factors, such as position of the consonant in the word and surrounding sounds [@lisker1977rapid]. <!-- NOTE: (?) why not "phonological contrast"? -->
Furthermore, correlates of the contrast can even be found away from the consonant, in temporally distant speech units.
For example, the initial /l/ of the English words *led* and *let* is affected by the voicing of the final consonant (/t, d/) [@hawkins2004influence].

@@

When looking at phrase-level temporal domains, the number of possible analytical pipelines increases exponentially. <!-- NOTE: (JVC) this is hyperbole, but seems (possibly) more appropriate than "explodes" -->
This plurality of analytical paths is illustrated in Figure \@ref(fig:forkingPaths).
<!-- When comparing two utterances such as (2a) and (2b), there are many things to consider.  NOTE: (JVC) (I think we can do without this sentence.) -->
Even if we have decided to compare fundamental frequency (f0) of only the word *another* across the two utterances, there are still many choices to be made, all of which need to be justified.
For example, we could measure f0 at specific points in time like the onset of the window, the offset, or the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarize f0 across the entire window and extract the mean, median or standard deviation of f0, all of which have been used to analyze speech data [see @gordon2017acoustic].
And the garden of forking paths does not stop here.
<!-- TODO: add examples of pitch tracking options -->
<!-- NOTE: (TR) is the figure better now? -->
Figure \@ref(fig:forkingPaths), illustrates a specific option to automatically calculate f0.
Knowing that the signals and the measures thereof can be noisy, another option would be to smooth them using one algorithm and degree over others, or remove values that substantially deviate from surrounding values or expectations, either manually or automatically, and so on. <!-- NOTE: (JVC) Im having a bit of trouble with this sentence. I tried to fix it and now I am realizing I don't quite follow (JVC) -->

@@

Even looking at morphosyntactically identical expressions, speakers can modulate utterances via suprasegmental acoustic properties like temporal and spectral modifications of the segments involved [e.g., @ladd2008intonational].
Most prominently, languages use intonation to signal discourse relationships between referents (among other functions).
Intonation marks discourse-relevant referents for being new or given information to guide the listeners’ interpretation of incoming messages.
Beyond structuring information relative to the discourse, a few studies suggested that speakers might use intonation to signal atypical lexical combinations [e.g. dimitrova2008prosodic; dimitrova2009did].
Referential expressions such as *blue banana* were produced with greater prosodic prominence than more typical referent such as *yellow banana*.
If true, these findings are in line with the idea of resource-efficient, rational language users who modulate their speech in order to facilitate listeners' comprehension process.
However, the original conclusions of above studies were based on small (n=10), unjustified sample sizes.
Moreover, the data is not available, and the statistical analysis used anti-conservative co-variance structures, leaving reason to doubt the original conclusions. <!-- NOTE: (JVC) "lending doubt on" sounded a bit weird to me (JVC) -->

@@

<!-- TODO: add materials link and mailing lists -->
The initiating authors (SC, JC, TR) created an online landing page providing a general description of the project [https://many-speech-analyses.github.io](https://many-speech-analyses.github.io) and a short pre-recorded slide show that summarizes the study and research question (LINK). <!-- NOTE: same as above, or OSF -->
The project will be advertised via social media, using mailing lists for linguistic and psychological societies, and via word of mouth.
The target population comprises active speech science researchers with a graduate/doctoral degree (or currently studying for a graduate/doctoral degree) in relevant disciplines.
Researchers can choose to work independently or in a small team.
For the sake of simplicity, we will refer to single researcher or small teams as "analysis teams".

@@

<!-- NOTE: I understand "posterior probability" to refer to a point estimate, i.e., the posterior probability that B > 0 = some value between 0 and 1. I may be mistaken, but I think you are referring to the posterior distribution in the above paragraph -->

@@

<!-- NOTE: I keep changing meta-analytical to meta-analytic (the -cal sounds weird to me). My unscientific google searching shows that people use both, but meta-analytic is much more common (nearly 4 million hits vs. 500k) -->
<!-- NOTE: (SC) I tend to use analytical to mean "related to analysis" and "analytic" in opposition to "synthetic" (from their use in logic and linguistics) but I am flex as long as we are consistent. -->

@@

### Analytical and demographic predictors affecting effect sizes {#ana-factors}
<!-- NOTE: (TR) I think we should use the term predictors. Factors is really ambiguous -->
As a second step, we will investigate (a) the extent to which the individual standardized effect sizes are affected by a series of predictors related to analytical and demographic factors (see below); and (b) the extent to which deviations from the meta-analytic point estimate by the individual standardized effect sizes relate to those same predictors.
We will fit two Bayesian regression models, one for (a) and one for (b).
<!-- TODO: check if stand. effect sizes and deviations are equivalent. -->
with each team's group-level coefficient $\alpha_{\text{t}[i]}$ as the outcome variable, and the analytical and demographic factors described below as predictors.
The coefficient $\alpha_{\text{t}[i]}$ can be interpreted as the deviation score of the team's $\eta_i$ from the meta-analytic point estimate, allowing us to determine potential effects of analytical and demographic factors on how the teams deviate from the meta-analytic mean.
Since each $\alpha_{\text{t}[i]}$ comes with uncertainty, quantified by its standard deviation $\sigma_{\alpha_{\text{t}[i]}}$, we will fit a measurement-error model, as we did in the meta-analytic model above.
